%!TEX root = ../thesis.tex
% ******************************* Thesis Appendix A ****************************
\chapter*{List of Publications}
\addcontentsline{toc}{chapter}{List of Publications}
\section*{International Journals}
%{enumi}
%\renewcommand{\labelenumi}{[IJ-\arabic{enumi}]}
%\begin{enumerate}{[IJ-\arabic{enumi}]}%[label={[RI-\arabic*]}]
\renewcommand{\labelenumi}{[IJ-\arabic{enumi}]}
\begin{etaremune}%[label={[RI-\arabic*]}]
	\item \textbf{Othman, R.}, Skrzypczak, A., Lou\"{e}t, Y. `` Reduced Complexity Detectors Based on SOQPSK Linearzations,'', in \emph{IEEE Transactions on Aerospace and Electronic Systems}, 2018 (Major Revision)
	\item \textbf{Othman, R.}, Skrzypczak, A., Lou\"{e}t, Y., `` PAM Decomposition of Ternary CPM With Duobinary Encoding,'', in \emph{IEEE Transactions on Communications}, vol. 65, no. 10, pp. 4274-4284, doi: 10.1109/TCOMM.2017.2723567, Oct 2017
\end{etaremune}

\section*{International Conferences}
%\renewcommand{\labelenumi}{[IC-\arabic{enumi}]}
\renewcommand{\labelenumi}{[IC-\arabic{enumi}]}
%\begin{enumerate}{[IC-\arabic{enumi}]}
\begin{etaremune}
	\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``Linear Decision Feedback Detector for SOQPSK,'', in \emph{URSI AP-RASC}, 2019 (Submitted)
	\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``A Reduced Complexity OQPSK-Type Detector For SOQPSK,'', in \emph{IEEE 88th Vehicular Technology Conference (VTC-Fall)},Chicago,Sep 2018 (to be published)
	\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``Joint Channel Estimation and Detection of SOQPSK Using the PAM Decomposition,'', in \emph{IEEE 25th International Conference on Telecommunications (ICT)}, St. Malo, pp. 154-158. doi: 10.1109/ICT.2018.8464828, Jun 2018 
\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``Analysis of Duobinary Encoding for CPM Signals,'', in \emph{URSI General Assembly and Symposium}, Montreal, Aug 2017
\item \textbf{Othman, R.}, Skrzypczak, A., Lou\"{e}t, Y. `` Phase offset estimation of SOQPSK waveform by the analysis of the angle distribution,'', in \emph{IEEE 13th International Symposium on Wireless Communication Systems}, Poznan, pp. 16-21. doi: 10.1109/ISWCS.2016.7600847, Sep 2016  
\end{etaremune}

\section*{National Conferences}
\renewcommand{\labelenumi}{[NC-\arabic{enumi}]}
\begin{etaremune}
\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``D\'{e}composition PAM d'un signal CPM avec encodage duo-binaire - Application \`{a} la modulation SOQPSK,'', in \emph{GRETSI}, Juan-Les-Pins, France, Sep 2017 
\end{etaremune}

\section*{Other National Communications}
\renewcommand{\labelenumi}{[ONC-\arabic{enumi}]}
\begin{etaremune}
\item	\textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``A New PAM Decomposition for SOQPSK in Aernautical Telemetry Context,'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jul 2017 
\item	\textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``Etude de techniques d’\'{e}galisation aveugles et aid\'{e}es pour les modulations de fr\'{e}quence,'', in \emph{SCEE Seminar}, Rennes, France, Jun 2017 
\item	\textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. ``Etude de techniques d’\'{e}galisation aveugles et aid\'{e}es pour les modulations de fr\'{e}quence,'', in \emph{Journ\'{e}e Des Doctorants de l'IETR}, Rennes, France, Jun 2016 	
\end{etaremune}

\section*{Patents}
\renewcommand{\labelenumi}{[PT-\arabic{enumi}]}
\begin{etaremune}
\item \textbf{Othman, R.}, Lou\"{e}t, Y., Skrzypczak, A. `` Nouvelle architecture de d\'{e}codage spatio-temporel bas\'{e}e sur la d\'{e}composition PAM pour la modulation SOQPSK-TG,'' (pending)
\end{etaremune}